### 2018-03-12 submission/manuscript.tex
disucssion read through

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

\usepackage{amsmath}

\usepackage{mathtools}

\begin{document}

\section{Making Sense of the Noise: Leveraging Existing 16S rRNA Gene
Surveys to Identify Key Community Members in Colorectal
Tumors}\label{making-sense-of-the-noise-leveraging-existing-16s-rrna-gene-surveys-to-identify-key-community-members-in-colorectal-tumors}

\begin{center}
\vspace{25mm}

Marc A Sze${^1}$ and Patrick D Schloss${^1}$${^\dagger}$

\vspace{20mm}

$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu

$1$ Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI

\end{center}

Co-author e-mails:

\begin{itemize}
\tightlist
\item
  \href{mailto:marcsze@med.umich.edu}{\nolinkurl{marcsze@med.umich.edu}}
\end{itemize}

\newpage

\linenumbers

\subsection{Abstract}\label{abstract}

\textbf{Background.} An increasing body of literature suggests that both
individual and collections of bacteria are associated with the
progression of colorectal cancer. As the number of studies investigating
these associations increases and the number of subjects in each study
increases, a meta-analysis to identify the associations that are the
most predictive of disease progression is warranted. For our
meta-analysis, we analyzed previously published 16S rRNA gene sequencing
data collected from feces (1737 individuals from 8 studies) and colon
tissue (492 total samples from 350 individuals from 7 studies).

\textbf{Results.} We quantified the odds ratios for individual bacterial
taxa that were associated with an individual having tumors relative to a
normal colon. Among the stool samples, there were no taxa that had a
significant odds ratio associated with adenoma and there were 8 taxa
with significant odds ratios (ORs) associated with carcinoma. Similarly,
among the tissue samples, there were no taxa that had a significant odds
ratio associated with adenoma and there were 3 taxa with significant
odds ratios associated with carcinoma. Among the significant odds
ratios, the association between individual taxa and tumor diagnosis was
equal or below 7.11. Because individual taxa had limited association
with tumor diagnosis, we trained Random Forest classification models
using only the taxa that had significant ORs, using the entire
collection of taxa found in each study, and using operational taxonomic
units defined based on a 97\% similarity threshold. All training
approaches yielded similar classification success as measured using the
Area Under the Curve. The ability to correctly classify individuals with
adenomas was poor and the ability to classify individuals with
carcinomas was considerably better using sequences from stool or tissue.

\newpage

\textbf{Importance.} This meta-analysis confirms previous results
indicating that individuals with adenomas cannot be readily classified
based on their bacterial community, but that those with carcinomas can.
Regardless of the dataset, we found a subset of the fecal community that
was associated with carcinomas was as predictive as the full community.

\subsubsection{Keywords}\label{keywords}

microbiota; colorectal cancer; polyps; adenoma; tumor; meta-analysis.

\newpage

\subsection{Background}\label{background}

Colorectal cancer (CRC) is a growing world-wide health problem in which
the microbiota has been hypothesized to have a role in disease
progression (1, 2). Numerous studies using murine models of CRC have
shown the importance of both individual microbes (3--7) and the overall
community (8--10) in tumorigenesis. Numerous case-control studies have
characterized the microbiota of individuals with colonic adenomas and
carcinomas in an attempt to identify biomarkers of disease progression
(6, 11--17). Because current CRC screening recommendations are poorly
adhered to due to socioeconomic status, test invasiveness, and frequency
of tests, development and validation of microbiome-associated biomarkers
for CRC progression could further attempts to develop non-invasive
diagnostics (18).

Recently, there has been an intense focus on identifying
microbiota-based biomarkers yielding a seemingly endless number of
candidate taxa. Some studies point towards mouth-associated genera such
as Fusobacterium, Peptostreptococcus, Parvimonas, and Porphyromonas that
are enriched in people with carcinomas (6, 11--17). Other studies have
identified members of \emph{Akkermansia}, \emph{Bacteroides},
\emph{Enterococcus}, \emph{Escherichia}, \emph{Klebsiella},
\emph{Mogibacterium}, \emph{Streptococcus}, and \emph{Providencia} are
also associated with carciomas (13--15). Additionally, \emph{Roseburia}
has been found in some studies to be more abundant in people with tumors
but in other studies it has been found to be either less abundant or no
different than what is found in subjects with normal colons (14, 17, 19,
20). There are strong results from tissue culture and murine models that
\emph{Fusobacterium nucleatum}, pks-positive strains of
\emph{Escherichia coli}, \emph{Streptococcus gallolyticus}, and an
entertoxin-producing strain of \emph{Bacteroides fragilis} are important
in the pathogenesis of CRC (5, 14, 21--24). These results point to a
causative role for the microbiota in CRC pathogenesis as well as their
potential as diagnostic biomarkers.

Most studies have focused on identifying biomarkers in patients with
carcinomas but there is a greater clinical need to identify biomarkers
associated with adenomas. Studies focusing on broad scale community
metrics have found that measures such as the total number of Operational
Taxonomic Units (OTUs) are decreased in those with adenomas versus
controls (25). Other studies have identified \emph{Acidovorax},
\emph{Bilophila}, \emph{Cloacibacterium}, \emph{Desulfovibrio},
\emph{Helicobacter}, \emph{Lactobacillus}, \emph{Lactococcus},
\emph{Mogibacterium}, and \emph{Pseudomonas} to be enriched in those
with adenomas (25--27). There are few genera that are enriched in
patients with adenoma or carcinoma tumors.

Confirming some of these previous findings, a recent meta-analysis found
that 16S rRNA gene sequences from members of \emph{Akkermansia},
\emph{Fusobacterium}, and \emph{Parvimonas} were fecal biomarkers for
the presence of carcinomas (28). Contrary to previous studies they found
sequences similar to members of \emph{Lactobacillus} and
\emph{Ruminococcus} to be enriched in patients with adenoma or carcinoma
relative to those with normal colons (12, 15, 16). In addition, they
found 16S rRNA gene sequences from members of \emph{Haemophilus},
\emph{Methanosphaera}, \emph{Prevotella}, \emph{Succinovibrio} were
enriched in patients with adenoma and \emph{Pantoea} were enriched in
patients with carcinomas. Although this meta-analysis was helpful for
distilling a large number of possible biomarkers, the aggregate number
of samples included in the analysis (n = 509) was smaller than several
larger case-control studies that have since been published (12, 27)

Here we provide an updated meta-analysis using 16S rRNA gene sequence
data from both feces (n = 1737) and colon tissue (492 samples from 350
individuals) from 14 studies (11--17, 19, 20, 23, 25--27, 29) {[}Table 1
\& 2{]}. We expand both the breadth and scope of the previous
meta-analysis to investigate whether biomarkers describing the bacterial
community or specific members of the community can more accurately
classify patients as having adenoma or carcinoma. Our results suggest
that the bacterial community changes as disease severity worsens and
that a subset of the microbial community can be used to diagnose the
presence of carcinoma.

\newpage

\subsection{Results}\label{results}

\textbf{\emph{Lower bacterial diversity is associated with increased
odds ratio (OR) of tumors.}} We first assessed whether variation in
broad community metrics like total number of operational taxonomic units
(OTUs) (i.e.~richness), the evenness of their abundance, and the overall
diversity was associated with disease stage after controlling for study
and variable region differences. In stool, there was a significant
decrease in both evenness and diversity as disease severity progressed
from normal to adenoma to carcinoma (P-value = 0.025 and 0.043,
respectively) {[}Figure 1{]}; there was no significant difference for
richness (P-value = 0.21). We next tested whether the decrease in these
community metrics translated into significant ORs for having an adenoma
or carcinoma. For fecal samples, the ORs for richness were not
significantly greater than 1.0 for adenoma or carcinoma (P-value = 0.40)
{[}Figure 2A{]}. The ORs for evenness were significantly higher than 1.0
for adenoma (OR = 1.3 (1.02 - 1.65)(95\% Confidence Interval), P-value =
0.035) and carcinoma (OR = 1.66 (1.2 - 2.3), P-value = 0.0021) {[}Figure
2B{]}. The ORs for diversity were only significantly greater than 1.0
for carcinoma (OR = 1.61 (1.14 - 2.28), P-value = 0.0069), but not for
adenoma (P-value = 0.11) {[}Figure 2C{]}. Although these OR are
significantly greater than 1.0, it is doubtful that these are clinically
meaningful values.

Similar to our analysis of sequences obtained from stool samples, we
repeated the analysis using sequences obtained from colon tissue. There
were no significant changes in richness, evenness, or diversity as
disease severity progressed from control to adenoma to carcinoma
(P-value \textgreater{} 0.05). We next analyzed the OR, for matched
(i.e.~where unaffected tissue and tumors were obtained from the same
individual) and unmatched (i.e.~where unaffected tissue and tumor tissue
were not obtained from the same individual) tissue samples. The ORs for
adenoma and carcinoma by any measure were not significantly different
from 1.0 (P-value \textgreater{} 0.05) {[}Figure S1 \& Table S1{]}. This
is likely due to the combination of a small effect size, as suggested
from the results using stool, and the relatively small number of studies
and size of studies used in the analysis.

\textbf{\emph{Disease progression is associated with community-wide
changes in composition and abundance.}} Based on the differences in
evenness and diversity, we next asked whether there were community-wide
differences in the structure of the communities associated with
different disease stages. We identified significant bacterial community
differences in the stool of patients with adenomas relative to those
with normal colons in 1 of 4 studies and in patients with carcinomas
relative to those with normal colons in 6 of 7 studies (PERMANOVA;
P-value \textless{} 0.05) {[}Table S2{]}. Similar to the analyses using
stool samples, there were significant differences in bacterial community
structure between subjects with normal colons and those with adenoma (1
of 2 studies) and carcinoma (1 of 3 studies) {[}Table S2{]}. For studies
that used matched samples no differences in bacterial community
structures were observed {[}Table S2{]}. Combined, these results
indicate that there are consistent and significant community-wide
changes in the fecal community structure of subjects with carcinomas.
However, the signal observed in subjects with adenomas or when using
tissue samples was not as consistent. This is likely due to a smaller
effect size or the relatively small sample sizes among the studies that
characterized the tissue microbiota.

\textbf{\emph{Individual taxa are associated with significant ORs for
carcinomas.}} Next we identified those taxa that were associated with
ORs that were significantly associated with having a normal colon or the
presence of adenomas or carcinomas. No taxa had a significant OR for the
presence of adenomas when we used data collected from stool or tissue
samples (Table S3 \& S4). In contrast, 8 taxa had significant ORs for
the presence of carcinomas using data from stool samples. Of these, 4
are commonly associated with the oral cavity: \emph{Fusobacterium} (OR =
2.74 (1.95 - 3.85)), \emph{Parvimonas} (OR = 3.07 (2.11 - 4.46)),
\emph{Porphyromonas} (OR = 3.2 (2.26 - 4.54)), and
\emph{Peptostreptococcus} (OR = 7.11 (3.84 - 13.17)) {[}Table S3{]}. The
other 4 were \emph{Clostridium XI} (OR = 0.65 (0.49 - 0.86)),
\emph{Enterobacteriaceae} (OR = 1.79 (1.33 - 2.41)), \emph{Escherichia}
(OR = 2.15 (1.57 - 2.95)), and \emph{Ruminococcus} (OR = 0.63 (0.48 -
0.83)). Among the data collected from tissue samples, only unmatched
carcinoma samples had taxa with a significant OR. Those included
\emph{Dorea} (OR = 0.35 (0.22 - 0.55)), \emph{Blautia} (OR = 0.47 (0.3 -
0.73)), and \emph{Weissella} (OR = 5.15 (2.02 - 13.14)).
Mouth-associated genera were not significantly associated with an
increased OR for carcinoma in tissue samples {[}Table S4{]}. For
example, \emph{Fusobacterium} had an OR of 3.98 (1.19 - 13.24; however,
due to the small number of studies and considerable variation in the
data, the Benjimani-Hochberg corrected P-value was 0.93 {[}Table S4{]}.
It is interesting to note that \emph{Ruminococcus} and members of
\emph{Clostridium XI} in stool and \emph{Dorea} and \emph{Blautia} in
tissue had ORs that were significantly less than 1.0, which suggests
that these populations are protective against the development of
carcinomas. Overall, there was no overlap in the taxa with significant
OR between stool and tissue samples.

\textbf{\emph{Individual taxa with a significant OR do a poor job of
differentiating subjects with normal colons and those with carcinoma.}}
We next asked whether those taxa that had a significant OR associated
with having a normal colon or carcinomas could be used individually, to
classify subjects as having a normal colon or carcinomas. OR values were
caluclated based on whether the relative abundance for a taxon in a
subject was above or below the median relative abundance for that taxon
across all subjects in a study. To measure the ability of these taxa to
classify individuals we instead generated receiver operator
characteristic (ROC) curves for each taxon in each study and calculated
the area under the curve (AUC). This allowed us to use a more fluid
relative abundance threshold for defining disease status. Using data
from stool samples, the 8 taxa did no better at classifying the subjects
than one would expect by chance (i.e.~AUC=0.50) {[}Figure 3A{]}. The
taxa that performed the best included \emph{Clostridium XI},
\emph{Ruminococcus}, and \emph{Escherichia}, however, these had median
AUC values less than 0.588. Likewise, in unmatched tissue samples the 3
taxa with significant OR taxa were marginally better than one would
expect by chance {[}Figure 3B{]}. The relative abundance of \emph{Dorea}
was the best predictor of carcinomas and its median AUC was only 0.62.
These results suggest that although these taxa are associated with a
decreased or increased OR for the presences of carcinomas, individually,
they do a poor job of classifying a subject's disease status.

\textbf{\emph{Combined taxa model classifies subjects better than using
individual taxa.}} Instead of attempting to classify subjects based on
individual taxa, next we combined information from the individual taxa
and evaluated the ability to classify a subject's disease status using
Random Forest models. For data from stool samples, the combined model
had an AUC of 0.75, which was significantly higher than any of the AUC
values for the individual taxa (P-value \textless{} 0.033). For the full
taxa models using stool, \emph{Bacteroides} and \emph{Lachnospiraceae}
were the most common taxa in the top 10\% mean decrease in accuracy
(MDA) across studies {[}Figure S2{]}. Of the 3 taxa with significant
ORs, all 3 were among the top 10\% most important taxa as measured by
mean decrease in accuracy, in at least one study. Within the significant
OR taxa only models the most important taxa across study for stool were
\emph{Ruminococcus} and \emph{Clostridium XI} based on mean decrease in
accuracy {[}Figure 4A{]}. Similarly, using data from the unmatched
tissue samples, the combined model had an AUC of 0.77, which was
significantly higher than the AUC values for \emph{Blautia} and
\emph{Weissella} (P-value \textless{} 0.037). For the full taxa models
using unmatched tissue, \emph{Lachnospiraceae}, \emph{Bacteroidaceae},
and \emph{Ruminococcaceae} were the most common taxa in the top 10\%
mean decrease in accuracy across studies {[}Figure S3{]}. For the
significant OR taxa unmatched tissue models both \emph{Dorea} and
\emph{Blautia} were the most important based on mean decrease in
accuracy {[}Figure 4B{]}. Pooling the information from the taxa with
significant ORs results in a model that outperforms classifications made
using individual taxa.

\textbf{\emph{Performance of models based on taxa relative abundance in
full community are better than those based on taxa with significant
ORs.}} Next, we asked whether a Random Forest classification model built
using all of the taxa found in the communities would outperform the
models generated using those taxa with a significant OR. Similar to our
inability to identify taxa associated with a significant OR for the
presence of adenomas, the median AUCs to classify subjects as having
normal colons or having adenomas using data from stool or tissue samples
were only marginally better than 0.5 for any study (median AUC = 0.549
(0.367 - 0.971)(min-max)) {[}Figure 5A \& S4A{]}. In contrast, the
models for classifying subjects as having normal colons or having
carcinomas using data from stool or tissue samples yielded AUC values
meaningfully higher than 0.5 {[}Figure 5B \& S4B-C{]}. When we compared
the models based on all of the taxa in a community to models based on
the taxa with significant ORs, the results were mixed. Using the data
from stool samples we found that the AUC for 6 of 7 studies increased by
an average of 14.8\%) and AUC for the Flemer study decreased by 0.54\%).
The overall improvement in performance was statistically significant
(mean = 12.61\%, one-tailed paired T-test; P-value = 0.005). Similarly,
using the data from unmatched tissue samples we found that the AUC of
studies increased by an average of 19.11\% when we used all of the taxa
rather than the 3 taxa with significant ORs (one-tailed paired T-test;
P-value = 0.03). Although the significant OR taxa models can classify
individuals with and without carcinoma tumors, they are still missing
taxa from the full community models that can increase the model
accuracy.

\textbf{\emph{Performance of models based on OTU relative abundance in
full community are not significantly better than those based on taxa
with significant ORs.}} The previous models were based on relative
abundance data where sequences were classified to coarse taxonomic
assignments (i.e.~typically genus or family level). To determine whether
model performance improved with finer scale classification, we assigned
sequences to operational taxonomic units (OTUs) where the similarity
among sequences within an OTU was more than 97\%. We again found that
classification models built using all of the sequence data for a
community did a poor job of differentiating between subjects with normal
colons and those with adenomas (median AUC: 0.53 (0.37- 0.56)), but did
a good job of differentiating between subjects with normal colons and
those with carcinomas (median AUC: 0.71 (0.5- 0.9)). The OTU-based
models performed similarly to those constructed using the taxa with
significant ORs (one-tailed paired T-test; P-value = 0.966) and those
using all taxa (one-tailed paired T-test; P-value = 0.146) {[}Figure
4{]}. Among the OTUs that had the highest mean decrease in accuracy for
the OTU-based models, we found that OTUs that affiliated with all of the
8 taxa that had a significant OR were within the top 10\% for at least
one study. This result was surprising as it indicated that a finer scale
classification of the sequences and thus a larger number of features to
select from, did not yield improved classification of the subjects.

\textbf{\emph{Generalizability of taxon-based models trained on one
dataset to the other datasets.}} Considering the good performance of the
Random Forest models using taxa with a significant OR and using all of
the taxa, we next asked how well the models would perform when given
data from a different subject cohort. For instance, if a model was
trained using data from the Ahn study, we wanted to know how well it
would perform using the data from the Baxter study. We found the models
trained using the taxa with a significant OR all had a higher median AUC
than the models trained using all of the taxa when tested on the other
datasets {[}Figure 6 \& S5{]}. As might be expected, the difference
between the performance of the modelling approaches appeared to vary
with the size of the training cohort (R\textsuperscript{2} = 0.66)
{[}Figure 6{]}. These data suggest that given a sufficient number of
subjects with normal colons and carcinomas, Random Forest models trained
using a small number of taxa can accurately classify individuals from a
different cohort.

\newpage

\subsection{Discussion}\label{discussion}

We performed a meta-analysis to identify and validate microbiome-base
biomarkers that could be used to classify individuals as having normal
colons or colonic tumors using fecal or tissue samples. To our surprise,
Random Forest classification models constructed to differentiate
individuals with normal colons from those with carcinomas using a subset
of the community performed well relative to models constructed using the
full communities. When we applied the models trained on each dataset to
the other datasets in our study, we found that the models trained using
the subset of the communities performed better than those using the full
communities. These models were trained using data in which sequences
were assigned to bacterial taxa using a classifier that typically
assigned sequences to the family or genus level. When we attempted to
improve the specificity of the classification by using an OTU-based
approach the resulting models performed as well as those constructed
using coarse taxonomic assignments. These results are significant
because they strengthen the growing literature indicating a role of the
microbiome in tumorigenesis (9) and as a potential tool as a
non-invasive diagnostic and for assessing risk of disease and recurrence
(12, 30).

These results suggested that fine scale classification of sequences into
OTUs does not improve our classification models. This has been suggested
in previous literature where shotgun metagenomic data did not perform
better than 16S rRNA gene sequencing data in classifying individuals
with normal colons and those with carcinomas (31). We hypothesize that
fine scale classification may not result in better classification
because distribution of microbiota between individuals is patchy. In
contrast, models using coarser taxonomic assignments will pool the fine
scale diversity, resulting in less patchiness and better classification.
Furthermore, the ability of models trained using a subset of the
community to outperform those using the full community when testing the
models on the other datasets may also be a product of the patchiness of
the human-associated microbiota. The models based on the 8 taxa that had
significant ORs used taxa that were found in every study and tended to
have higher relative abundances. Similar to the OTU-based models, those
models based on the full community taxonomy assignments were still
sensitive to the patchy distribution of taxa. Regardless, it is
encouraging that a collection of 8 taxa could reliably classify
individuals as having carcinomas considering the differences in cohorts,
DNA extraction procedures, regions of the 16S rRNA gene, and sequencing
methods.

When used separately to classify individuals with carcinomas, the taxa
with significant ORs could not reliably classify individuals {[}Figure
3{]}. This result further supports the hypothesis that
carcinoma-associated microbiota have a patchy distribution. Two
individuals may have had the same classification, based on the relative
abundance of different populations within this group of 8 taxa. Although
these results only reflect associations with disease, it is tempting to
hypothesize that the patchiness represents distinct mechanisms of
exacerbating tumorigenesis or that multiple taxa have the same mechanism
of exacerbating tumorigenesis. For example, strains of \emph{Escherichia
coli} and \emph{Fusobacterium nucleatum} have been shown to worsen
inflammation in mouse models of tumorigenesis (5, 6, 21). In contrast to
the patchiness of the taxa that were positively associated with
carcinomas, potentially beneficial taxa had a more consistent
association {[}Figure 6{]}. This result was particularly interesting
because members of these taxa (i.e. \emph{Ruminococcus} and
\emph{Clostridium XI} in stool and \emph{Dorea} and \emph{Blautia} in
tissue) are thought to be beneficial due to their involvement in
production of anti-inflammatory short chain fatty acids (32--34).

All of the adenoma classification models performed poorly and is not
inconsistent with previous studies (27, 30). However, the classification
results are at odds with results of the multitarget microbiota test
(MMT) from Baxter, \emph{et al.} (12) who observed an AUC of 0.755 when
applied to individuals with adenomas. There are two major differences
between the models generated in this meta-analysis and that analysis.
The MMT attempted to classify individuals as having a normal colon or
having colonic lesions (i.e.~adenomas or carcinomas) and not adenomas
alone. Further, the MMT incorporated fecal immunoglobulin test (FIT)
data while our models only used 16S rRNA gene sequencing data. Because
FIT data were not available for the other studies in our meta-analysis,
it was not possible to validate the MMT approach. The ability to
differentiate between individuals with and without adenomas is an
important problem since early detection of tumors is critical. However,
it is possible that we might have been able to detect differences in the
bacterial community if individuals with non-advanced and advanced
adenomas were separated. This is a clinically relevant distinction since
advanced adenomas are at highest risk of progressing to a carcinoma. The
initial changes of the microbiota during tumorigenesis could be focal to
where the initial adenoma develops and would not be easily assessed
using fecal samples from an individual with non-advanced adenomas.
Unfortunately, distinguishing between individuals with advanced and
non-advanced adenomas was not possible in our meta-analysis since the
studies did not provide the clinical data needed to make that
distinction.

Stool samples represent a non-invasive approach to assessing the
structure of the gut microbiota and are potentially useful for
diagnosing individuals as having colonic tumors. However, they do not
reflect the structure of the mucosal microbiota (35). Regardless, the
taxa that were the most important in the stool-based models overlapped
with those from the models trained using the data from unmatched and
matched colon tissue samples {[}Figure S3{]}. Mucosal biopsies are
preferred for focused mechanistic studies and have offered researchers
the opportunity to sample healthy and diseased tissue from the same
individuals (i.e.~matched) using each individual as their own control or
in a cross sectional design (i.e.~unmatched). Because obtaining these
samples is invasive, carries risks to the individual, and is expensive,
studies investigating the structure of the mucosal microbiota generally
have a limited number of participants. Thus, it was not surprising that
tissue-based studies did not provide clearer associations between the
mucosal microbiota and the presence of tumors. Interestingly,
Fusobacterium, which has received increased recent attention for its
potential role in tumorigenesis (6) was not consistently identified
across the studies in our meta-analysis. This could be due to the
relatively small number of individuals in the limited number of studies.
The classification models trained using the tissue-based data performed
well when tested with the training data (Figure S4), but performed
poorly when tested on the other tissue-associated datasets (Figure S5).
Disturbingly, taxa that are commonly associated with reagent
contamination (e.g. \emph{Novosphingobium}, \emph{Acidobacteria Gp2},
\emph{Sphingomonas}, etc.) were detected within the tissue datasets.
Such contamination is common in studies where there is relatively low
bacterial biomass (36). The lack of replication among the tissue-based
biomarkers may be a product of the relatively small number of studies
and individuals per study and possible reagent contamination.

Among our stool samples, we failed to identify several notable
populations that are commonly associated with carcinomas including an
enterotoxigenic strain of \emph{Bacteroides fragilis} (ETBF) and
\emph{Streptococcus gallolyticus} subsp. \emph{gallolyticus} (22, 24).
ETBF have been found in tumors in the proximal colon where they tend to
form biofilms (20, 37). Considering DNA from bacteria that are more
prevalent in the proximal colon may be degraded by the time it leaves
the body, it is not surprising that we failed to identify a significant
OR for Bacteroides with carcinomas. In addition, since our approach
could only classify sequences to the genus level and there are likely
multiple Bacteroides populations in the colon, it is possible that
sequences from ETBF and non-oncogenic Bacteroides were pooled. This
would then reduce the OR between Bacterioides and whether an individual
had carcinomas. It is also necessary to distinguish between populations
that are biomarkers for a disease and those that are known to cause
disease. The former may also have a causative role in the disease.
Although the latter have been shown to have a causative role, they may
appear at low relative abundance, be found in specific locations, or may
have a highly patchy distribution among affected individuals.

Meta-analyses are a useful tool in microbiome research because they can
demonstrate whether a result can be replicated and facilitate new
discoveries by pooling multiple independent investigations. There have
been several meta-analyses similar to this study that have sought
biomarkers for obesity (38--40), inflammatory bowel disease (39), and
colorectal cancer (28). Considering microbiome research is particularly
prone to hype and overgeneralization of results, these analyses are
critical. For example, previous meta-analyses have demonstrated that
there are no clear fecal biomarkers for obesity (39, 40). Such
meta-analyses are difficult to perform because the underlying 16S rRNA
gene sequence data are not publicly available, metadata are missing,
incomplete, or vague, sequence data are of poor quality or derived by
non-standard approaches, and the original studies were significantly
underpowered. Reluctance to publish negative results (i.e.~the ``file
drawer effect'') is also likely to skew our understanding of the
relationship between microbiota and disease. Better attention to these
specifics will increase the reproducibility and replicability of
microbiota studies and make it easier to perform these crucial
meta-analyses. Moving forward, meta-analyses will be important tools to
help aggregate and find commonalities across studies when investigating
the microbiota in the context of a specific disease (28, 38--40).

Our meta-analysis suggests a strong association between the gut
microbiota and colon tumorigenesis. By aggregating the results from
studies that sequenced the 16S rRNA gene from fecal and tissue samples,
we are able to provide evidence supporting the use of microbial
biomarkers to diagnose the presence of colonic tumors. Further
development of microbial biomarkers should focus on including other
biomarkers (e.g.~FIT), better categorizing of people with adenomas, and
expanding datasets to include larger numbers of individuals. Based on
prior research into the physiology of the biomarkers we identified, it
is likely that they have a causative role in tumorigenesis. Their patchy
distribution across individuals suggests that there are either multiple
mechanisms causing disease or a single mechanism (e.g.~inflammation)
that can be mediated by multiple, diverse bacteria.

\newpage

\subsection{Methods}\label{methods}

\textbf{\emph{Datasets.}} The studies used for this meta-analysis were
identified through the review articles written by Keku, et al. (41) and
Vogtmann, et al. (42). Additional studies, not mentioned in those
reviews were obtained based on the authors' knowledge of the literature.
Studies were included that used tissue or feces as their sample source
for 454 or Illumina 16S rRNA gene sequencing. Some studies (N = 12) were
excluded because they did not have publicly available sequences, did not
use 454 or Illumina machines, or did not have metadata in which the
authors were able to share. We were able to obtain sequence data and
metadata from the following studies: Ahn, \emph{et al.} (11), Baxter,
\emph{et al.} (12), Brim, \emph{et al.} (29), Burns, \emph{et al.} (15),
Chen, \emph{et al.} (13), Dejea, \emph{et al.} (20), Flemer, \emph{et
al.} (17), Geng, \emph{et al.} (19), Hale, \emph{et al.} (27), Kostic,
\emph{et al.} (43), Lu, \emph{et al.} (26), Sanapareddy, \emph{et al.}
(25), Wang, \emph{et al.} (14), Weir, \emph{et al.} (23), and Zeller,
\emph{et al.} (16). The Zackular (44) study was excluded because their
90 individuals were contained within the larger Baxter study (12). The
Kostic study was excluded because after we processed the sequences, all
of the case samples had 100 or fewer sequences. The final analysis
included 14 studies (Tables 1 and 2). There were seven studies with only
fecal samples (Ahn, Baxter, Brim, Hale, Wang, Weir, and Zeller), five
studies with only tissue samples (Burns, Dejea, Geng, Lu, Sanapareddy),
and two studies with both fecal and tissue samples (Chen and Flemer).
After curating the sequences, 1737 stool samples and 492 tissue samples
remained in the analysis {[}Tables 1 and 2{]}.

\textbf{\emph{Sequence Processing.}} Raw sequence data and metadata were
primarily obtained from the Sequence Read Archive (SRA) and dbGaP. Other
sequence and metadata were obtained directly from the authors (n = 4,
(17, 23, 25, 27)). Each dataset was processed separately using mothur
(v1.39.3) (45) using the default quality filtering methods for both 454
and Illumina sequence data. If it was not possible to use the defaults
because the sequences were trimmed too much, then the stated quality
cut-offs from the original study were used. Chimeric sequences were
identified and removed using VSEARCH (46). The curated sequences were
assigned to OTUs at 97\% similarity using the OptiClust algorithm (47)
and classified to the deepest taxonomic level that had 80\% support
using the naïve Bayesian classifier trained on the RDP taxonomy outline
(version 14, (48)).

\textbf{\emph{Community analysis.}} We calculated alpha diversity
metrics (i.e.~OTU richness, evenness, and Shannon diversity) for each
sample. Within each dataset, we ensured that the data followed a normal
distribution using power transformations. Using the transformed data, we
tested the hypothesis that individuals with normal colons, adenomas, and
carcinomas had significantly different alpha diversity metrics using
linear mixed-effect models. We also calculated the OR for each study and
metric by considering any value above the median alpha diversity value
to be positive. We measured the dissimilarity between individuals by
calculating the pairwise Bray-Curtis index and used PERMANOVA (49) to
test whether individuals with normal colons were significantly different
from those with adenomas or carcinomas. Finally, after binning sequences
into the deepest taxa that the naïve Bayesian classifier could calssify
the sequences, we quantified the ORs for individuals having an adenoma
or carcinoma and corrected for multiple comparisons using the
Benjamini-Hochberg method (50). Again, for each taxon, if the relative
abundance was greater than the median relative abundance for that taxon
in the study, the individual was considered to be positive.

\textbf{\emph{Random Forest classification analysis.}} To classify
individuals as having normal colons or tumors, we built Random Forest
classification models for each dataset and comparison using taxa with
significant ORs (after multiple comparison correction), all taxa, or
OTUs. Because no taxa were identified as having a significant OR
associated with adenomas using stool samples or tissue samples,
classification models based on OR data were not constructed to classify
individuals as having normal colons or adenomas. Within the training
dataset, 10-fold cross validation (5-fold cross validation for small
datasets) was used to build a model that was then evaluated on the
testing set. For the models constructed using the taxa with significant
ORs, the default mtry setting was used to train the model and this model
was tested on the other datasets in the meta-analysis. The reported AUC
values are the AUCs for the application of the model on the test sets.
For the OTU-based models, the dataset was split into training (80\% of
samples) and testing (20\%) sets and 10-fold cross validation (5-fold
cross validation for small datasets) on the training set was used to
generate the model for the testing set. The original 80/20 split and
fitting was repeated 100 times and the average AUC from these 100
repeats was reported. The Mean Decrease in Accuracy (MDA), a measure of
the importance of each taxon to the overall model was used to rank the
taxa used in each model. For all models, the default setting used was
\(\sqrt{p}\), where p is all the variables used in the respective model.
Normally, \(\sqrt{p}\) has been found to be what is chosen as the ideal
mtry after model tuneing (51).

\textbf{\emph{Statistical Analysis.}} All statistical analysis after
sequence processing utilized the R (v3.4.3) software package (52). For
OTU richness, evenness, and Shannon diversity analysis, values were
power transformed using the rcompanion (v1.11.1) package (53) and then
Z-score normalized using the car (v2.1.6) package (54). Testing for OTU
richness, evenness, and Shannon diversity differences utilized linear
mixed-effect models created using the lme4 (v1.1.15) package (55) to
correct for study, repeat sampling of individuals (tissue only), and 16S
hyper-variable region used. Odds ratios (OR) were analyzed using both
the epiR (v0.9.93) and metafor (v2.0.0) packages (56, 57) by assessing
how many individuals with and without disease were above and below the
overall median value within each specific study. OR significance testing
utilized the chi-squared test. Diversity differences measured by the
Bray-Curtis index utilized the creation of distance matrix and testing
with PERMANOVA executed with the vegan (v2.4.5) package (58). Random
Forest models were built using both the caret (v6.0.78) and randomForest
(v4.6.12) packages (59, 60). All figures were created using both ggplot2
(v2.2.1) and gridExtra (v2.3) packages (61, 62).

\textbf{\emph{Reproducible Methods.}} The code and analysis can be found
at
\url{https://github.com/SchlossLab/Sze_CRCMetaAnalysis_Microbiome_2017}.
Unless otherwise mentioned, the accession number of raw sequences from
the studies used in this analysis can be found directly in the
respective batch file in the GitHub repository or in the original
manuscript.

\newpage

\subsubsection{Acknowledgements}\label{acknowledgements}

The authors would like to thank all the study participants who were a
part of each of the individual studies analyzed. We would also like to
thank each of the study authors for making their sequencing reads and
metadata available for use. Finally, we would like to thank the members
of the Schloss lab for their valuable feedback and proofreading during
the formulation of this manuscript.

\newpage

\subsection{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-siegel_cancer_2016}{}
1. \textbf{Siegel, R. L.}, \textbf{K. D. Miller}, and \textbf{A. Jemal}.
2016. Cancer statistics, 2016. CA: a cancer journal for clinicians
\textbf{66}:7--30.

\hypertarget{ref-flynn_metabolic_2016}{}
2. \textbf{Flynn, K. J.}, \textbf{N. T. Baxter}, and \textbf{P. D.
Schloss}. 2016. Metabolic and Community Synergy of Oral Bacteria in
Colorectal Cancer. mSphere \textbf{1}.

\hypertarget{ref-goodwin_polyamine_2011}{}
3. \textbf{Goodwin, A. C.}, \textbf{C. E. Destefano Shields}, \textbf{S.
Wu}, \textbf{D. L. Huso}, \textbf{X. Wu}, \textbf{T. R. Murray-Stewart},
\textbf{A. Hacker-Prietz}, \textbf{S. Rabizadeh}, \textbf{P. M. Woster},
\textbf{C. L. Sears}, and \textbf{R. A. Casero}. 2011. Polyamine
catabolism contributes to enterotoxigenic Bacteroides fragilis-induced
colon tumorigenesis. Proceedings of the National Academy of Sciences of
the United States of America \textbf{108}:15354--15359.

\hypertarget{ref-abed_fap2_2016}{}
4. \textbf{Abed, J.}, \textbf{J. E. M. Emgård}, \textbf{G. Zamir},
\textbf{M. Faroja}, \textbf{G. Almogy}, \textbf{A. Grenov}, \textbf{A.
Sol}, \textbf{R. Naor}, \textbf{E. Pikarsky}, \textbf{K. A. Atlan},
\textbf{A. Mellul}, \textbf{S. Chaushu}, \textbf{A. L. Manson},
\textbf{A. M. Earl}, \textbf{N. Ou}, \textbf{C. A. Brennan}, \textbf{W.
S. Garrett}, and \textbf{G. Bachrach}. 2016. Fap2 Mediates Fusobacterium
nucleatum Colorectal Adenocarcinoma Enrichment by Binding to
Tumor-Expressed Gal-GalNAc. Cell Host \& Microbe \textbf{20}:215--225.

\hypertarget{ref-arthur_intestinal_2012}{}
5. \textbf{Arthur, J. C.}, \textbf{E. Perez-Chanona}, \textbf{M.
Mühlbauer}, \textbf{S. Tomkovich}, \textbf{J. M. Uronis}, \textbf{T.-J.
Fan}, \textbf{B. J. Campbell}, \textbf{T. Abujamel}, \textbf{B. Dogan},
\textbf{A. B. Rogers}, \textbf{J. M. Rhodes}, \textbf{A. Stintzi},
\textbf{K. W. Simpson}, \textbf{J. J. Hansen}, \textbf{T. O. Keku},
\textbf{A. A. Fodor}, and \textbf{C. Jobin}. 2012. Intestinal
inflammation targets cancer-inducing activity of the microbiota. Science
(New York, N.Y.) \textbf{338}:120--123.

\hypertarget{ref-kostic_fusobacterium_2013}{}
6. \textbf{Kostic, A. D.}, \textbf{E. Chun}, \textbf{L. Robertson},
\textbf{J. N. Glickman}, \textbf{C. A. Gallini}, \textbf{M. Michaud},
\textbf{T. E. Clancy}, \textbf{D. C. Chung}, \textbf{P. Lochhead},
\textbf{G. L. Hold}, \textbf{E. M. El-Omar}, \textbf{D. Brenner},
\textbf{C. S. Fuchs}, \textbf{M. Meyerson}, and \textbf{W. S. Garrett}.
2013. Fusobacterium nucleatum potentiates intestinal tumorigenesis and
modulates the tumor-immune microenvironment. Cell Host \& Microbe
\textbf{14}:207--215.

\hypertarget{ref-wu_human_2009}{}
7. \textbf{Wu, S.}, \textbf{K.-J. Rhee}, \textbf{E. Albesiano},
\textbf{S. Rabizadeh}, \textbf{X. Wu}, \textbf{H.-R. Yen}, \textbf{D. L.
Huso}, \textbf{F. L. Brancati}, \textbf{E. Wick}, \textbf{F.
McAllister}, \textbf{F. Housseau}, \textbf{D. M. Pardoll}, and
\textbf{C. L. Sears}. 2009. A human colonic commensal promotes colon
tumorigenesis via activation of T helper type 17 T cell responses.
Nature Medicine \textbf{15}:1016--1022.

\hypertarget{ref-zackular_manipulation_2016}{}
8. \textbf{Zackular, J. P.}, \textbf{N. T. Baxter}, \textbf{G. Y. Chen},
and \textbf{P. D. Schloss}. 2016. Manipulation of the Gut Microbiota
Reveals Role in Colon Tumorigenesis. mSphere \textbf{1}.

\hypertarget{ref-zackular_gut_2013}{}
9. \textbf{Zackular, J. P.}, \textbf{N. T. Baxter}, \textbf{K. D.
Iverson}, \textbf{W. D. Sadler}, \textbf{J. F. Petrosino}, \textbf{G. Y.
Chen}, and \textbf{P. D. Schloss}. 2013. The gut microbiome modulates
colon tumorigenesis. mBio \textbf{4}:e00692--00613.

\hypertarget{ref-baxter_structure_2014}{}
10. \textbf{Baxter, N. T.}, \textbf{J. P. Zackular}, \textbf{G. Y.
Chen}, and \textbf{P. D. Schloss}. 2014. Structure of the gut microbiome
following colonization with human feces determines colonic tumor burden.
Microbiome \textbf{2}:20.

\hypertarget{ref-ahn_human_2013}{}
11. \textbf{Ahn, J.}, \textbf{R. Sinha}, \textbf{Z. Pei}, \textbf{C.
Dominianni}, \textbf{J. Wu}, \textbf{J. Shi}, \textbf{J. J. Goedert},
\textbf{R. B. Hayes}, and \textbf{L. Yang}. 2013. Human gut microbiome
and risk for colorectal cancer. Journal of the National Cancer Institute
\textbf{105}:1907--1911.

\hypertarget{ref-baxter_microbiota-based_2016}{}
12. \textbf{Baxter, N. T.}, \textbf{M. T. Ruffin}, \textbf{M. A. M.
Rogers}, and \textbf{P. D. Schloss}. 2016. Microbiota-based model
improves the sensitivity of fecal immunochemical test for detecting
colonic lesions. Genome Medicine \textbf{8}:37.

\hypertarget{ref-chen_human_2012}{}
13. \textbf{Chen, W.}, \textbf{F. Liu}, \textbf{Z. Ling}, \textbf{X.
Tong}, and \textbf{C. Xiang}. 2012. Human intestinal lumen and
mucosa-associated microbiota in patients with colorectal cancer. PloS
One \textbf{7}:e39743.

\hypertarget{ref-wang_structural_2012}{}
14. \textbf{Wang, T.}, \textbf{G. Cai}, \textbf{Y. Qiu}, \textbf{N.
Fei}, \textbf{M. Zhang}, \textbf{X. Pang}, \textbf{W. Jia}, \textbf{S.
Cai}, and \textbf{L. Zhao}. 2012. Structural segregation of gut
microbiota between colorectal cancer patients and healthy volunteers.
The ISME journal \textbf{6}:320--329.

\hypertarget{ref-burns_virulence_2015}{}
15. \textbf{Burns, M. B.}, \textbf{J. Lynch}, \textbf{T. K. Starr},
\textbf{D. Knights}, and \textbf{R. Blekhman}. 2015. Virulence genes are
a signature of the microbiome in the colorectal tumor microenvironment.
Genome Medicine \textbf{7}:55.

\hypertarget{ref-zeller_potential_2014}{}
16. \textbf{Zeller, G.}, \textbf{J. Tap}, \textbf{A. Y. Voigt},
\textbf{S. Sunagawa}, \textbf{J. R. Kultima}, \textbf{P. I. Costea},
\textbf{A. Amiot}, \textbf{J. Böhm}, \textbf{F. Brunetti}, \textbf{N.
Habermann}, \textbf{R. Hercog}, \textbf{M. Koch}, \textbf{A. Luciani},
\textbf{D. R. Mende}, \textbf{M. A. Schneider}, \textbf{P.
Schrotz-King}, \textbf{C. Tournigand}, \textbf{J. Tran Van Nhieu},
\textbf{T. Yamada}, \textbf{J. Zimmermann}, \textbf{V. Benes},
\textbf{M. Kloor}, \textbf{C. M. Ulrich}, \textbf{M. von Knebel
Doeberitz}, \textbf{I. Sobhani}, and \textbf{P. Bork}. 2014. Potential
of fecal microbiota for early-stage detection of colorectal cancer.
Molecular Systems Biology \textbf{10}:766.

\hypertarget{ref-flemer_tumour-associated_2017}{}
17. \textbf{Flemer, B.}, \textbf{D. B. Lynch}, \textbf{J. M. R. Brown},
\textbf{I. B. Jeffery}, \textbf{F. J. Ryan}, \textbf{M. J. Claesson},
\textbf{M. O'Riordain}, \textbf{F. Shanahan}, and \textbf{P. W.
O'Toole}. 2017. Tumour-associated and non-tumour-associated microbiota
in colorectal cancer. Gut \textbf{66}:633--643.

\hypertarget{ref-GimenoGarca2012}{}
18. \textbf{García, A. Z. G.} 2012. Factors influencing colorectal
cancer screening participation. Gastroenterology Research and Practice.
Hindawi Limited \textbf{2012}:1--8.

\hypertarget{ref-geng_diversified_2013}{}
19. \textbf{Geng, J.}, \textbf{H. Fan}, \textbf{X. Tang}, \textbf{H.
Zhai}, and \textbf{Z. Zhang}. 2013. Diversified pattern of the human
colorectal cancer microbiome. Gut Pathogens \textbf{5}:2.

\hypertarget{ref-dejea_microbiota_2014}{}
20. \textbf{Dejea, C. M.}, \textbf{E. C. Wick}, \textbf{E. M.
Hechenbleikner}, \textbf{J. R. White}, \textbf{J. L. Mark Welch},
\textbf{B. J. Rossetti}, \textbf{S. N. Peterson}, \textbf{E. C.
Snesrud}, \textbf{G. G. Borisy}, \textbf{M. Lazarev}, \textbf{E. Stein},
\textbf{J. Vadivelu}, \textbf{A. C. Roslani}, \textbf{A. A. Malik},
\textbf{J. W. Wanyiri}, \textbf{K. L. Goh}, \textbf{I. Thevambiga},
\textbf{K. Fu}, \textbf{F. Wan}, \textbf{N. Llosa}, \textbf{F.
Housseau}, \textbf{K. Romans}, \textbf{X. Wu}, \textbf{F. M.
McAllister}, \textbf{S. Wu}, \textbf{B. Vogelstein}, \textbf{K. W.
Kinzler}, \textbf{D. M. Pardoll}, and \textbf{C. L. Sears}. 2014.
Microbiota organization is a distinct feature of proximal colorectal
cancers. Proceedings of the National Academy of Sciences of the United
States of America \textbf{111}:18321--18326.

\hypertarget{ref-ecoli_Arthur_2014}{}
21. \textbf{Arthur, J. C.}, \textbf{R. Z. Gharaibeh}, \textbf{M.
Mühlbauer}, \textbf{E. Perez-Chanona}, \textbf{J. M. Uronis}, \textbf{J.
McCafferty}, \textbf{A. A. Fodor}, and \textbf{C. Jobin}. 2014.
Microbial genomic analysis reveals the essential role of inflammation in
bacteria-induced colorectal cancer. Nature Communications. Springer
Nature \textbf{5}:4724.

\hypertarget{ref-strep_Aymeric_2017}{}
22. \textbf{Aymeric, L.}, \textbf{F. Donnadieu}, \textbf{C. Mulet},
\textbf{L. du Merle}, \textbf{G. Nigro}, \textbf{A. Saffarian},
\textbf{M. Bérard}, \textbf{C. Poyart}, \textbf{S. Robine}, \textbf{B.
Regnault}, \textbf{P. Trieu-Cuot}, \textbf{P. J. Sansonetti}, and
\textbf{S. Dramsi}. 2017. Colorectal cancer specific conditions
promoteStreptococcus gallolyticusgut colonization. Proceedings of the
National Academy of Sciences. Proceedings of the National Academy of
Sciences \textbf{115}:E283--E291.

\hypertarget{ref-weir_stool_2013}{}
23. \textbf{Weir, T. L.}, \textbf{D. K. Manter}, \textbf{A. M. Sheflin},
\textbf{B. A. Barnett}, \textbf{A. L. Heuberger}, and \textbf{E. P.
Ryan}. 2013. Stool microbiome and metabolome differences between
colorectal cancer patients and healthy adults. PloS One
\textbf{8}:e70803.

\hypertarget{ref-bfrag_Boleij_2014}{}
24. \textbf{Boleij, A.}, \textbf{E. M. Hechenbleikner}, \textbf{A. C.
Goodwin}, \textbf{R. Badani}, \textbf{E. M. Stein}, \textbf{M. G.
Lazarev}, \textbf{B. Ellis}, \textbf{K. C. Carroll}, \textbf{E.
Albesiano}, \textbf{E. C. Wick}, \textbf{E. A. Platz}, \textbf{D. M.
Pardoll}, and \textbf{C. L. Sears}. 2014. The bacteroides fragilis toxin
gene is prevalent in the colon mucosa of colorectal cancer patients.
Clinical Infectious Diseases. Oxford University Press (OUP)
\textbf{60}:208--215.

\hypertarget{ref-sanapareddy_increased_2012}{}
25. \textbf{Sanapareddy, N.}, \textbf{R. M. Legge}, \textbf{B. Jovov},
\textbf{A. McCoy}, \textbf{L. Burcal}, \textbf{F. Araujo-Perez},
\textbf{T. A. Randall}, \textbf{J. Galanko}, \textbf{A. Benson},
\textbf{R. S. Sandler}, \textbf{J. F. Rawls}, \textbf{Z. Abdo},
\textbf{A. A. Fodor}, and \textbf{T. O. Keku}. 2012. Increased rectal
microbial richness is associated with the presence of colorectal
adenomas in humans. The ISME journal \textbf{6}:1858--1868.

\hypertarget{ref-lu_mucosal_2016}{}
26. \textbf{Lu, Y.}, \textbf{J. Chen}, \textbf{J. Zheng}, \textbf{G.
Hu}, \textbf{J. Wang}, \textbf{C. Huang}, \textbf{L. Lou}, \textbf{X.
Wang}, and \textbf{Y. Zeng}. 2016. Mucosal adherent bacterial dysbiosis
in patients with colorectal adenomas. Scientific Reports
\textbf{6}:26337.

\hypertarget{ref-hale_shifts_2017}{}
27. \textbf{Hale, V. L.}, \textbf{J. Chen}, \textbf{S. Johnson},
\textbf{S. C. Harrington}, \textbf{T. C. Yab}, \textbf{T. C. Smyrk},
\textbf{H. Nelson}, \textbf{L. A. Boardman}, \textbf{B. R. Druliner},
\textbf{T. R. Levin}, \textbf{D. K. Rex}, \textbf{D. J. Ahnen},
\textbf{P. Lance}, \textbf{D. A. Ahlquist}, and \textbf{N. Chia}. 2017.
Shifts in the Fecal Microbiota Associated with Adenomatous Polyps.
Cancer Epidemiology, Biomarkers \& Prevention: A Publication of the
American Association for Cancer Research, Cosponsored by the American
Society of Preventive Oncology \textbf{26}:85--94.

\hypertarget{ref-shah_leveraging_2017}{}
28. \textbf{Shah, M. S.}, \textbf{T. Z. DeSantis}, \textbf{T.
Weinmaier}, \textbf{P. J. McMurdie}, \textbf{J. L. Cope}, \textbf{A.
Altrichter}, \textbf{J.-M. Yamal}, and \textbf{E. B. Hollister}. 2017.
Leveraging sequence-based faecal microbial community survey data to
identify a composite biomarker for colorectal cancer. Gut.

\hypertarget{ref-brim_microbiome_2013}{}
29. \textbf{Brim, H.}, \textbf{S. Yooseph}, \textbf{E. G. Zoetendal},
\textbf{E. Lee}, \textbf{M. Torralbo}, \textbf{A. O. Laiyemo},
\textbf{B. Shokrani}, \textbf{K. Nelson}, and \textbf{H. Ashktorab}.
2013. Microbiome analysis of stool samples from African Americans with
colon polyps. PloS One \textbf{8}:e81352.

\hypertarget{ref-Sze2017}{}
30. \textbf{Sze, M. A.}, \textbf{N. T. Baxter}, \textbf{M. T. Ruffin},
\textbf{M. A. M. Rogers}, and \textbf{P. D. Schloss}. 2017.
Normalization of the microbiota in patients after treatment for colonic
lesions. Microbiome. Springer Nature \textbf{5}.

\hypertarget{ref-Hannigan2017}{}
31. \textbf{Hannigan, G. D.}, \textbf{M. B. Duhaime}, \textbf{M. T.
Ruffin}, \textbf{C. C. Koumpouras}, and \textbf{P. D. Schloss}. 2017.
Diagnostic potential \& the interactive dynamics of the colorectal
cancer virome. Cold Spring Harbor Laboratory.

\hypertarget{ref-Venkataraman2016}{}
32. \textbf{Venkataraman, A.}, \textbf{J. R. Sieber}, \textbf{A. W.
Schmidt}, \textbf{C. Waldron}, \textbf{K. R. Theis}, and \textbf{T. M.
Schmidt}. 2016. Variable responses of human microbiomes to dietary
supplementation with resistant starch. Microbiome. Springer Nature
\textbf{4}.

\hypertarget{ref-Herrmann2018}{}
33. \textbf{Herrmann, E.}, \textbf{W. Young}, \textbf{V.
Reichert-Grimm}, \textbf{S. Weis}, \textbf{C. Riedel}, \textbf{D.
Rosendale}, \textbf{H. Stoklosinski}, \textbf{M. Hunt}, and \textbf{M.
Egert}. 2018. In vivo assessment of resistant starch degradation by the
caecal microbiota of mice using RNA-based stable isotope probingA
proof-of-principle study. Nutrients. MDPI AG \textbf{10}:179.

\hypertarget{ref-Reichardt2017}{}
34. \textbf{Reichardt, N.}, \textbf{M. Vollmer}, \textbf{G. Holtrop},
\textbf{F. M. Farquharson}, \textbf{D. Wefers}, \textbf{M. Bunzel},
\textbf{S. H. Duncan}, \textbf{J. E. Drew}, \textbf{L. M. Williams},
\textbf{G. Milligan}, \textbf{T. Preston}, \textbf{D. Morrison},
\textbf{H. J. Flint}, and \textbf{P. Louis}. 2017. Specific
substrate-driven changes in human faecal microbiota composition contrast
with functional redundancy in short-chain fatty acid production. The
ISME Journal. Springer Nature \textbf{12}:610--622.

\hypertarget{ref-Flynn_preprint_2017}{}
35. \textbf{Flynn, K. J.}, \textbf{M. T. Ruffin}, \textbf{D. K.
Turgeon}, and \textbf{P. D. Schloss}. 2017. Spatial variation of the
native colon microbiota in healthy adults. Cold Spring Harbor
Laboratory.

\hypertarget{ref-Salter_contamination_2014}{}
36. \textbf{Salter, S. J.}, \textbf{M. J. Cox}, \textbf{E. M. Turek},
\textbf{S. T. Calus}, \textbf{W. O. Cookson}, \textbf{M. F. Moffatt},
\textbf{P. Turner}, \textbf{J. Parkhill}, \textbf{N. J. Loman}, and
\textbf{A. W. Walker}. 2014. Reagent and laboratory contamination can
critically impact sequence-based microbiome analyses. BMC Biology.
Springer Nature \textbf{12}.

\hypertarget{ref-Purcell2017}{}
37. \textbf{Purcell, R. V.}, \textbf{J. Pearson}, \textbf{A. Aitchison},
\textbf{L. Dixon}, \textbf{F. A. Frizelle}, and \textbf{J. I. Keenan}.
2017. Colonization with enterotoxigenic bacteroides fragilis is
associated with early-stage colorectal neoplasia. PLOS ONE. Public
Library of Science (PLoS) \textbf{12}:e0171602.

\hypertarget{ref-Sze2016}{}
38. \textbf{Sze, M. A.}, and \textbf{P. D. Schloss}. 2016. Looking for a
signal in the noise: Revisiting obesity and the microbiome. mBio.
American Society for Microbiology \textbf{7}:e01018--16.

\hypertarget{ref-Walters2014}{}
39. \textbf{Walters, W. A.}, \textbf{Z. Xu}, and \textbf{R. Knight}.
2014. Meta-analyses of human gut microbes associated with obesity and
IBD. FEBS Letters. Wiley-Blackwell \textbf{588}:4223--4233.

\hypertarget{ref-Finucane2014}{}
40. \textbf{Finucane, M. M.}, \textbf{T. J. Sharpton}, \textbf{T. J.
Laurent}, and \textbf{K. S. Pollard}. 2014. A taxonomic signature of
obesity in the microbiome? Getting to the guts of the matter. PLoS ONE.
Public Library of Science (PLoS) \textbf{9}:e84689.

\hypertarget{ref-keku_gastrointestinal_2015}{}
41. \textbf{Keku, T. O.}, \textbf{S. Dulal}, \textbf{A. Deveaux},
\textbf{B. Jovov}, and \textbf{X. Han}. 2015. The gastrointestinal
microbiota and colorectal cancer. American Journal of Physiology -
Gastrointestinal and Liver Physiology \textbf{308}:G351--G363.

\hypertarget{ref-vogtmann_epidemiologic_2016}{}
42. \textbf{Vogtmann, E.}, and \textbf{J. J. Goedert}. 2016.
Epidemiologic studies of the human microbiome and cancer. British
Journal of Cancer \textbf{114}:237--242.

\hypertarget{ref-kostic_genomic_2012}{}
43. \textbf{Kostic, A. D.}, \textbf{D. Gevers}, \textbf{C. S.
Pedamallu}, \textbf{M. Michaud}, \textbf{F. Duke}, \textbf{A. M. Earl},
\textbf{A. I. Ojesina}, \textbf{J. Jung}, \textbf{A. J. Bass},
\textbf{J. Tabernero}, \textbf{J. Baselga}, \textbf{C. Liu}, \textbf{R.
A. Shivdasani}, \textbf{S. Ogino}, \textbf{B. W. Birren}, \textbf{C.
Huttenhower}, \textbf{W. S. Garrett}, and \textbf{M. Meyerson}. 2012.
Genomic analysis identifies association of Fusobacterium with colorectal
carcinoma. Genome Research \textbf{22}:292--298.

\hypertarget{ref-zackular_human_2014}{}
44. \textbf{Zackular, J. P.}, \textbf{M. A. M. Rogers}, \textbf{M. T.
Ruffin}, and \textbf{P. D. Schloss}. 2014. The human gut microbiome as a
screening tool for colorectal cancer. Cancer Prevention Research
(Philadelphia, Pa.) \textbf{7}:1112--1121.

\hypertarget{ref-schloss_introducing_2009}{}
45. \textbf{Schloss, P. D.}, \textbf{S. L. Westcott}, \textbf{T.
Ryabin}, \textbf{J. R. Hall}, \textbf{M. Hartmann}, \textbf{E. B.
Hollister}, \textbf{R. A. Lesniewski}, \textbf{B. B. Oakley}, \textbf{D.
H. Parks}, \textbf{C. J. Robinson}, \textbf{J. W. Sahl}, \textbf{B.
Stres}, \textbf{G. G. Thallinger}, \textbf{D. J. Van Horn}, and
\textbf{C. F. Weber}. 2009. Introducing mothur: Open-Source,
Platform-Independent, Community-Supported Software for Describing and
Comparing Microbial Communities. Appl.Environ.Microbiol.
\textbf{75}:7537--7541.

\hypertarget{ref-rognes_vsearch_2016}{}
46. \textbf{Rognes, T.}, \textbf{T. Flouri}, \textbf{B. Nichols},
\textbf{C. Quince}, and \textbf{F. Mahé}. 2016. VSEARCH: A versatile
open source tool for metagenomics. PeerJ \textbf{4}:e2584.

\hypertarget{ref-westcott_opticlust_2017}{}
47. \textbf{Westcott, S. L.}, and \textbf{P. D. Schloss}. 2017.
OptiClust, an Improved Method for Assigning Amplicon-Based Sequence Data
to Operational Taxonomic Units. mSphere \textbf{2}.

\hypertarget{ref-rdp_Wang2007}{}
48. \textbf{Wang, Q.}, \textbf{G. M. Garrity}, \textbf{J. M. Tiedje},
and \textbf{J. R. Cole}. 2007. Naive bayesian classifier for rapid
assignment of rRNA sequences into the new bacterial taxonomy. Applied
and Environmental Microbiology. American Society for Microbiology
\textbf{73}:5261--5267.

\hypertarget{ref-permanova_Anderson2013}{}
49. \textbf{Anderson, M. J.}, and \textbf{D. C. I. Walsh}. 2013.
PERMANOVA, ANOSIM, and the mantel test in the face of heterogeneous
dispersions: What null hypothesis are you testing? Ecological
Monographs. Wiley-Blackwell \textbf{83}:557--574.

\hypertarget{ref-benjamini_controlling_1995}{}
50. \textbf{Benjamini, Y.}, and \textbf{Y. Hochberg}. 1995. Controlling
the false discovery rate: A practical and powerful approach to multiple
testing. Journal of the Royal Statistical Society. Series B
(Methodological) \textbf{57}:289--300.

\hypertarget{ref-Breiman2001}{}
51. \textbf{Breiman, L.} 2001. Machine Learning. Springer Nature
\textbf{45}:5--32.

\hypertarget{ref-r_citation_2017}{}
52. \textbf{R Core Team}. 2017. R: A language and environment for
statistical computing. R Foundation for Statistical Computing, Vienna,
Austria.

\hypertarget{ref-rcompanion_citation_2017}{}
53. \textbf{Mangiafico, S.} 2017. Rcompanion: Functions to support
extension education program evaluation.

\hypertarget{ref-car_citation_2011}{}
54. \textbf{Fox, J.}, and \textbf{S. Weisberg}. 2011. An R companion to
applied regressionSecond. Sage, Thousand Oaks CA.

\hypertarget{ref-lme4_citation_2015}{}
55. \textbf{Bates, D.}, \textbf{M. Mächler}, \textbf{B. Bolker}, and
\textbf{S. Walker}. 2015. Fitting linear mixed-effects models using
lme4. Journal of Statistical Software \textbf{67}:1--48.

\hypertarget{ref-epir_citation_2017}{}
56. \textbf{Telmo Nunes, M. S. with contributions from}, \textbf{C.
Heuer}, \textbf{J. Marshall}, \textbf{J. Sanchez}, \textbf{R. Thornton},
\textbf{J. Reiczigel}, \textbf{J. Robison-Cox}, \textbf{P. Sebastiani},
\textbf{P. Solymos}, \textbf{K. Yoshida}, \textbf{G. Jones}, \textbf{S.
Pirikahu}, \textbf{S. Firestone}, and \textbf{R. Kyle.} 2017. EpiR:
Tools for the analysis of epidemiological data.

\hypertarget{ref-metafor_citation_2010}{}
57. \textbf{Viechtbauer, W.} 2010. Conducting meta-analyses in R with
the metafor package. Journal of Statistical Software \textbf{36}:1--48.

\hypertarget{ref-vegan_citation_2017}{}
58. \textbf{Oksanen, J.}, \textbf{F. G. Blanchet}, \textbf{M. Friendly},
\textbf{R. Kindt}, \textbf{P. Legendre}, \textbf{D. McGlinn}, \textbf{P.
R. Minchin}, \textbf{R. B. O'Hara}, \textbf{G. L. Simpson}, \textbf{P.
Solymos}, \textbf{M. H. H. Stevens}, \textbf{E. Szoecs}, and \textbf{H.
Wagner}. 2017. Vegan: Community ecology package.

\hypertarget{ref-caret_citation_2017}{}
59. \textbf{Jed Wing, M. K. C. from}, \textbf{S. Weston}, \textbf{A.
Williams}, \textbf{C. Keefer}, \textbf{A. Engelhardt}, \textbf{T.
Cooper}, \textbf{Z. Mayer}, \textbf{B. Kenkel}, \textbf{the R Core
Team}, \textbf{M. Benesty}, \textbf{R. Lescarbeau}, \textbf{A. Ziem},
\textbf{L. Scrucca}, \textbf{Y. Tang}, \textbf{C. Candan}, and
\textbf{T. Hunt.} 2017. Caret: Classification and regression training.

\hypertarget{ref-randomforest_citation_2002}{}
60. \textbf{Liaw, A.}, and \textbf{M. Wiener}. 2002. Classification and
regression by randomForest. R News \textbf{2}:18--22.

\hypertarget{ref-ggplot2_citation_2009}{}
61. \textbf{Wickham, H.} 2009. Ggplot2: Elegant graphics for data
analysis. Springer-Verlag New York.

\hypertarget{ref-gridextra_citation_2017}{}
62. \textbf{Auguie, B.} 2017. GridExtra: Miscellaneous functions for
``grid'' graphics.

\newpage

\textbf{Table 1: Total Individuals in each Study Included in the Stool
Analysis}

\footnotesize

\begin{longtable}[]{@{}cccccc@{}}
\toprule
Study & Data Stored & Region & Control (n) & Adenoma (n) & Carcinoma
(n)\tabularnewline
\midrule
\endhead
Ahn & DBGap & V3-4 & 148 & 0 & 62\tabularnewline
Baxter & SRA & V4 & 172 & 198 & 120\tabularnewline
Brim & SRA & V1-3 & 6 & 6 & 0\tabularnewline
Flemer & Author & V3-4 & 37 & 0 & 43\tabularnewline
Hale & Author & V3-5 & 473 & 214 & 17\tabularnewline
Wang & SRA & V3 & 56 & 0 & 46\tabularnewline
Weir & Author & V4 & 4 & 0 & 7\tabularnewline
Zeller & SRA & V4 & 50 & 37 & 41\tabularnewline
\bottomrule
\end{longtable}

\normalsize
\newpage

\textbf{Table 2: Studies with Tissue Samples Included in the Analysis}

\footnotesize

\begin{longtable}[]{@{}cccccc@{}}
\toprule
Study & Data Stored & Region & Control (n) & Adenoma (n) & Carcinoma
(n)\tabularnewline
\midrule
\endhead
Burns & SRA & V5-6 & 18 & 0 & 16\tabularnewline
Chen & SRA & V1-3 & 9 & 0 & 9\tabularnewline
Dejea & SRA & V3-5 & 31 & 0 & 32\tabularnewline
Flemer & Author & V3-4 & 103 & 37 & 94\tabularnewline
Geng & SRA & V1-2 & 16 & 0 & 16\tabularnewline
Lu & SRA & V3-4 & 20 & 20 & 0\tabularnewline
Sanapareddy & Author & V1-2 & 38 & 0 & 33\tabularnewline
\bottomrule
\end{longtable}

\normalsize
\newpage

\textbf{Figure 1: Significant Bacterial Community Metrics for Adenoma or
Carcinoma in Stool.} A) Adenoma evenness. B) Carcinoma evenness. C)
Carcinoma Shannon diversity. Blue represents controls and red represents
either adenoma (panel A) or carcinoma (panel B and C). The black lines
represent the median value for each repsective group.

\textbf{Figure 2: Odds Ratio for Adenoma or Carcinoma based on Bacterial
Community Metrics in Stool.} A) Community-based odds ratio for adenoma.
B) Community-based odds ratio for carcinoma. Colors represent the
different variable regions used within the respective study.

\textbf{Figure 3: The AUC of Indivdiual Significant OR Taxa to classify
Carcinoma.} A) Stool samples. B) Unmatched tissue samples. The larger
circle represents the median AUC of all studies and the smaller circles
represent the individual AUC for a particular study. The dotted line
denotes an AUC of 0.5.

\textbf{Figure 4: Most Important Members in Significant OR Taxa
Carcinoma Models.} A) Common taxa in the top 10 percent for carcinoma
Random Forest stool-based models. B) Common taxa in the top 10 percent
for carcinoma Random Forest unmatched tissue-based models. Blue
represents less important and red represents more important to the
Random Forest Model. White means that particular taxa was not in the top
10\%.

\textbf{Figure 5: Stool Random Forest Model Train AUCs.} A) Adenoma
random forest model AUCs between all genera, all OTU, and select model
based on significant OR taxa. B) Carcinoma random forest model AUCs
between all genera, all OTU, and select model based on significant OR
taxa. The black line represents the median AUC for the respective group.
If no values are present in the singificant OR taxa group then there
were no significant taxa identified and no model was tested.

\textbf{Figure 6: Stool Random Forest Genus-Based Model Test AUCs.} A)
Test AUCs of adenoma models using all genera across study. B) Test AUCs
of carcinoma models using all genera or significant OR taxa only. The
black line represents the AUC at 0.5. The red lines represent the median
AUC of all test AUCs for a specific study.

\newpage

\textbf{Figure S1: Odds Ratio for Adenoma or Carcinoma based on
Bacterial Community Metrics in Tissue.} A) Community-based odds ratio
for adenoma. B) Community-based odds ratio for carcinoma. Colors
represent the different variable regions used within the respective
study.

\textbf{Figure S2: Most Common Taxa Across Carcinoma Full Community
Stool Study Models.} A) Common taxa in the top 10 percent for carcinoma
Random Forest all taxa-based models. B) Common taxa in the top 10
percent for carcinoma Random Forest all OTU-based models. Blue
represents less important and red represents more important to the
Random Forest Model. White means that particular taxa was not in the top
10\%.

\textbf{Figure S3: Most Common Genera Across Full Community Tissue Study
Models.} A) Common genera in the top 10 percent for matched carcinoma
Random Forest all genera-based models. B) Common genera in the top 10
percent for unmatched carcinoma Random Forest all genera-based models.
C) Common genera in the top 10 percent for matched carcinoma Random
Forest all OTU-based models. D) Common genera in the top 10 percent for
unmatched carcinoma Random Forest all OTU-based models. Blue represents
less important and red represents more important to the Random Forest
Model. White means that particular taxa was not in the top 10\%.

\textbf{Figure S4: Tissue Random Forest Model Train AUCs.} A) Adenoma
random forest model AUCs between all genera, all OTU, and select model
based on significant OR taxa in unmatched and matched tissue. B)
Carcinoma random forest model AUCs between all genera, all OTU, and
select model based on significant OR taxa in unmatched and matched
tissue. The black line represents the median AUC for the respective
group. If no values are present in the singificant OR taxa group then
there were no significant taxa identified and no model was tested.

\textbf{Figure S5: Tissue Random Forest Genus-Based Model Test AUCs.} A)
Test AUCs of adenoma models using all genera across study. B) Test AUCs
of carcinoma models using all genera for matched tissue studies. C) Test
AUCs of carcinoma models using all genera or significant OR taxa only
for unmatched tissue studies The black line represents the AUC at 0.5.
The red lines represent the median AUC of all test AUCs for a specific
study.

\newpage

\end{document}